---
title: "Practical courses"
author: "Amin Madoui and Romuald Laso-Jadart"
date: "April 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
  # Load read counts file, filtering
  
```{r }
  source ("popgene_functions.R")  
  
 data=read.table('data/RC/OsimilisRC.txt',row.names = 1) 

 poplist = c("155","158","178","206","208","209","210")

 hist(data$GA206 + data$GB206,breaks = 100)

 data = FilteringCov(data,poplist,dev=2, minCov=5, maxCov = 150)
 
 hist(data$GA206 + data$GB206,breaks = 100)
 # Create an allele frequency matrix
    
freq = CreateFreqMatrix(data,poplist)

# Example of two loci with unwanted frequencies
barplot(freq["13564462",],ylim=c(0,1))
barplot(freq["10789206",],ylim=c(0,1))
 
 data = FilteringFreqExpression(data,poplist,GAF=0.1, TAF=0.05)
  
  
```
  #   Estimation of the genomic differentiation
  ##    Allele frequency spectrum

```{r }
 Gfreq = data$GB206/(data$GA206+data$GB206)  
 
 hist(Gfreq, breaks = 10, main="B allele frequency distribution")  
```
 
 ##  Pairwise-FST
 
```{r}
source ("popgene_functions.R")  
 
freq = CreateFreqMatrix(data,poplist) 

colnames(freq) = c("TARA_155","TARA_158","TARA_178","TARA_206","TARA_208","TARA_209","TARA_210")   

pairwiseFST = pwFst(freq)  

print(pairwiseFST)  
 
plotFST(pairwiseFST)  
```
 
 ## Isolation by distances
 
```{r}

euclideans = read.table("data/Euclidean_Arctic.txt", row.names=1,header = TRUE)
 
MantelPlot(euclideans,pairwiseFST)
 
mantel.euclideans = mantel(vegdist(pairwiseFST,na.rm = T),vegdist(euclideans,na.rm = T))

print(mantel.euclideans)
```

## Isolation by currents

```{r }
lagrangians = read.table("data/Lagrangian_Arctic.txt", row.names=1,header = TRUE)
 
MantelPlot(lagrangians,pairwiseFST)
 
mantel.lagrangians = mantel(vegdist(pairwiseFST,na.rm = T),vegdist(lagrangians,na.rm = T))

print(mantel.lagrangians)


```
  
 # Detection of loci under natural selection
 
 ## The LK outliers method 
 
```{r }

freq = CreateFreqMatrix(data,poplist) 

colnames(freq) = c("TARA_155","TARA_158","TARA_178","TARA_206","TARA_208","TARA_209","TARA_210")  

LK = LK(freq)  

head( LK)  

h = hist ( LK$LK , plot = F, breaks = 50 )  

n_pop = ncol(freq)-1  

hist ( LK$LK, freq = F, xlab = "LK" , breaks = 50 , main = "LK distribution")  

# add the neutral evolution  

curve ( dchisq( x, n_pop ) , lwd=3, col="orange" , add=T )  

# selected loci with high LK value 

selection_LK = LK[LK$q_value<0.1,]  

dim(selection_LK)  
```
 
 ## using the pcadapt package
```{r}
library(pcadapt)
 
freq = as.data.frame(CreateFreqMatrix(data,poplist)) 

colnames(freq) = c("TARA_155","TARA_158","TARA_178","TARA_206","TARA_208","TARA_209","TARA_210")   

freq_forPCA = read.pcadapt(t(freq), type = "pool")

pcadapt_result <- pcadapt(freq_forPCA)

plot(pcadapt_result, option = "scores", pop = colnames(freq))

plot(pcadapt_result, option = "screeplot")

plot(pcadapt_result, option = "qqplot")

q_values = p.adjust(pcadapt_result$pvalues, method = "BH")

freq$pcadapt_qval = q_values

freq[is.na(freq)] <- 1

dim(freq[freq$pcadapt_qval<0.05,])

freq$LK_qval = LK$q_value

dim(freq[freq$pcadapt_qval<0.05 & freq$LK_qval<0.1,])

```
 
 # Population-scale allele expression  
 
 ## Comparison of allele expression to genomic abundance
 
```{r }

 Gfreq = data$GB206/(data$GA206+data$GB206)  
 
 Tfreq = data$TB206/(data$TA206+data$TB206) 
 
 plot(Gfreq,Tfreq)  
 
 summary(lm(Tfreq~Gfreq-1))  
```
 
 ## Detection of allele-specific expression (ASE)
 ## Example with TARA_206
```{r}
 
ASE_206 = data[,c("GA206","GB206","TA206","TB206")]

ASE_206 = ASE_206[ASE_206[,1] >= 1 & ASE_206[,2] >= 1,]

Fisher_pvalue = apply(ASE_206,1,function (x) fisher.test (matrix(x, nrow = 2))$p.value) 

Fisher_qvalue = p.adjust(Fisher_pvalue, method="BH")  
 
ASE_206$Fisher_qvalue = Fisher_qvalue

dim(ASE_206[ASE_206$Fisher_qvalue<0.05,])

```
 
 ## Link between natural selection and allele specific expression
 
```{r}

selection_ASE_S206 = intersect(rownames(freq[freq$pcadapt_qval<0.05,]),
                        rownames(ASE_206[ASE_206$Fisher_qvalue < 0.05,])) 

phyper(length(selection_ASE_S206),   
       nrow(freq[freq$pcadapt_qval<0.05,]),  
       nrow(freq)-nrow(freq[freq$pcadapt_qval<0.05,]),  
       length(Fisher_qvalue[Fisher_qvalue<0.05]),
       lower.tail = F)  
  
```
 
 # Reference-free approach and metavariant species
 
```{r }
  
 MVS = readMvs("data/MVS/Ciliophora/")
  
```
 ## Genomic differentiation of MVS
 
```{r }
  
  plotFST(pairwiseFST = MVS@pwFst)
  
```
 
 ## Role of the environmental factors
 
```{r }
EnvTable = read.table("data/WorldOceanAtlas_data.txt",header = T)
envlist = list(Temperature = EnvTable[,c(1,7)],
             Salinity = EnvTable[,c(1,8)],
             Silicate = EnvTable[,c(1,4)],
             Po4 = EnvTable[,c(1,6)], 
             No3 = EnvTable[,c(1,5)])

LagrangianMatrix = read.table("data/Lagrangian_Matrix.txt",header = T)

ggplot(data = melt(as.matrix(LagrangianMatrix)), aes(Var1,Var2, fill = value)) + 
  geom_tile(color = "white",names ="Travel Time (days)") + scale_fill_viridis() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        axis.text.x = element_text(angle=90,vjust = 0.8 ), 
        axis.text.y = element_text(vjust = 0.8 )) +
  xlab("") + ylab("")
 
```
  
 ## Variance partitioning of environmental factors
  
```{r}
   
VariancePartition.Res = LinearMixedModelMVS(MVS,LagrangianMatrix,envlist)

PlotLMM(VariancePartition.Res) 

```
