---
title: "Practical courses"
author: "Amin Madoui and Romuald Laso-Jadart"
date: "26 février 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#   Estimation of the genomic differentiation
##    Allele frequency spectrum

 ```
 data=read.table('RC/S155.txt')
 colnames(data) = c("GA","GB","TA","TB")
 Gfreq = data$GB/(data$GA+data$GB)
 hist(Gfreq, breaks = 10, main="B allele frequency distribution")
 ```
 ##Pairwise-FST
 
 
 ##Isolation by currents2    
 # Detection of loci under natural selection
 ##FSToutliers method
 ##pcadapt method
 #Population-scale allele expression
 ##Comparison of allele expression to genomic abundance
 ##Detection of allele-specific expression
 ##Link between natural selection and allele specific ex-pression
 #Reference-free approach and metavariant species
 ##Genomic differentiation of MVS
 ##Role of the environmental factors
 ##Variance partitioning of environmental factors


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
